---
title: "Cohort Analysis"
author: "Jason Whittle"
date: "5/15/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(tidyverse)
data <- read.csv("DTC_NSC_20180514.csv")
og_data <- read.csv("mtl_dtc.csv")
data <- data %>% left_join(og_data, by = c("Requester.Return.Field" = "id"))
data$tech_grad <- data$graduated
data$ID <- data$Requester.Return.Field

data <- data %>% select("First.Name", "Last.Name", "ID", 
                        "Record.Found.Y.N", "College.Name", "X2.year...4.year", 
                        "Public...Private", "Enrollment.Begin", "Enrollment.End", 
                        "Enrollment.Status", "Graduated.", "Graduation.Date",
                        "Degree.Title", "Degree.Major.1", "College.Sequence",
                        "inst", "StartDate", "StopDate", "Birthdate", "tech_grad")
```

```{r, include=F}
library(lubridate)
Sys.setenv(TZ="America/Los_Angeles")
Sys.getenv("TZ")
data$Enrollment.Begin <- ymd(data$Enrollment.Begin)
data$Enrollment.End <- ymd(data$Enrollment.End)
data$Graduation.Date <- ymd(data$Graduation.Date)
data$Birthdate <- ymd(data$Birthdate)
data$StartDate <- ymd(data$StartDate)
data$StopDate <- ymd(data$StopDate)
data <- data %>% filter(College.Name != "DAVIS TECHNICAL COLLEGE") # Mountain land is not in the data. 
```

# Data summary

```{r, include=F}
data %>% group_by(inst, tech_grad) %>% summarise(n = n_distinct(ID))

```


# Cohort analysis

```{r, include=F}
data$twoyear <- data$StartDate %m+% months(24)
data$fouryear <- data$StartDate %m+% months(48)
data$sixyear <- data$StartDate %m+% months(72)
```


```{r, include=F}

# all graduates
data %>% summarise(n = n_distinct(ID)) # 2103
data %>% filter(Graduated. == "Y") %>% summarise(n = n_distinct(ID)) # 241 (11.4%)
data %>% filter(Graduated. == "Y") %>% filter(Graduation.Date < twoyear) %>% summarise(n = n_distinct(ID)) #138 (6.5%)
data %>% filter(Graduated. == "Y") %>% filter(Graduation.Date < fouryear) %>% summarise(n = n_distinct(ID)) #180 (8.5%)
data %>% filter(Graduated. == "Y") %>% filter(Graduation.Date < sixyear) %>% summarise(n = n_distinct(ID)) # 213 (10.1)
```



```{r, include=F}
# MTL  
data %>% filter(inst =="mtl") %>% summarise(n = n_distinct(ID)) # 436
data %>% filter(Graduated. == "Y") %>% filter(inst =="mtl") %>% filter(Graduation.Date < twoyear) %>% summarise(n = n_distinct(ID)) # 12 (2.7%)
data %>% filter(Graduated. == "Y") %>% filter(inst =="mtl") %>% filter(Graduation.Date < fouryear) %>% summarise(n = n_distinct(ID)) # 15 (3.4%)
data %>% filter(Graduated. == "Y") %>% filter(inst =="mtl") %>% filter(Graduation.Date < sixyear) %>% summarise(n = n_distinct(ID)) # 19 (4.6%)
data %>% filter(Graduated. == "Y") %>% filter(inst =="mtl") %>% summarise(n = n_distinct(ID)) # 24 (5.5%)
```

```{r}
mtl <- c(436, 12, 15, 19, 24)
mtl_pct <- c("-", 2.7, 3.4, 4.6, 5.5)
mtl <- rbind(mtl, mtl_pct)
colnames(mtl) <- c("Total", "Two year", "Four year", "Six year", "All grads")
row.names(mtl) <- c("Students", "Percent")
knitr::kable(mtl, caption = "Mountain Land College cohort graduations")
```

```{r, include=F}
# DTC
data %>% filter(inst =="dtc") %>% summarise(n = n_distinct(ID)) # 1667
data %>% filter(Graduated. == "Y") %>% filter(inst =="dtc") %>% filter(Graduation.Date < twoyear) %>% summarise(n = n_distinct(ID)) # 126 (7.6%)
data %>% filter(Graduated. == "Y") %>% filter(inst =="dtc") %>% filter(Graduation.Date < fouryear) %>% summarise(n = n_distinct(ID)) # 165 (9.9%)
data %>% filter(Graduated. == "Y") %>% filter(inst =="dtc") %>% filter(Graduation.Date < sixyear) %>% summarise(n = n_distinct(ID)) # 194 (11.6%)
data %>% filter(Graduated. == "Y") %>% filter(inst =="dtc") %>% summarise(n = n_distinct(ID)) # 217 (13%)
```

```{r}
dtc <- c(1667, 126, 165, 194, 217)
dtc_pct <- c("-", 7.6, 9.9, 11.6, 13)
dtc <- rbind(dtc, dtc_pct)
colnames(dtc) <- c("Total", "Two year", "Four year", "Six year", "All grads")
row.names(dtc) <- c("Students", "Percent")
knitr::kable(dtc, caption = "Davis Technical College cohort graduations")
```

# Transfer Institutions

Where do students transfer after starting at either Davis Technical College or Mountainland Technical College by year. These are unique student enrollments by institutions. If a student enrolled at SLCC and also at UVU they would be counted twice.

```{r, include=T}
# dtc
dtc_tot <- data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "dtc") %>% group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n))
dtc_2 <- data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "dtc") %>% filter(Enrollment.Begin < twoyear) %>%  group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n)) 
dtc_4 <- data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "dtc") %>% filter(Enrollment.Begin < fouryear) %>%  group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n))
dtc_6 <- data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "dtc") %>% filter(Enrollment.Begin < sixyear) %>%  group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n))

dtc_trns <- dtc_tot %>% left_join(dtc_2, by = "College.Name") %>% left_join(dtc_4, by = "College.Name") %>% left_join(dtc_6, by = "College.Name")
knitr::kable(dtc_trns[1:10,], 
             col.names = c("Institution Name", "Total", "W/in 2 years", "W/in 4 years", "W/in 6 years"),
             caption = "Transfer institition by time from stop data. Davis Tech")
```

```{r, include=T}
#mtl
mtl_tot <- data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "mtl") %>% group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n))
mtl_2 <- data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "mtl") %>% filter(Enrollment.Begin < twoyear) %>%  group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n))
mtl_4 <- data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "mtl") %>% filter(Enrollment.Begin < fouryear) %>%  group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n))
mtl_6 <- data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "mtl") %>% filter(Enrollment.Begin < sixyear) %>%  group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n))

mtl_trns <- mtl_tot %>% left_join(mtl_2, by = "College.Name") %>% left_join(mtl_4, by = "College.Name") %>% left_join(mtl_6, by = "College.Name")
knitr::kable(mtl_trns[1:10,], 
             col.names = c("Institution Name", "Total", "W/in 2 years", "W/in 4 years", "W/in 6 years"),
             caption = "Transfer institition by time from stop data. Mountain Land Tech")
```


```{r}
transfer_inst <- data %>% filter(Record.Found.Y.N == "Y") %>% 
  filter(Enrollment.Begin < twoyear) %>%  
  group_by(College.Name, inst) %>% 
  summarise(n = n_distinct(ID)) %>% 
  spread(inst, n)


```


```{r, include=F}
# mtl
data %>% filter(Record.Found.Y.N == "Y") %>% filter(inst == "mtl") %>% group_by(College.Name) %>% summarise(n = n_distinct(ID)) %>% arrange(desc(n))
```



## By year cohort graduation

Completions at other higher ed. institutions following an enrollment at DTC or MTC. Years since first enrollment at DTC or MTC.

```{r, include=F}
data$cohort <- year(data$StopDate)

all_student_cohort <- data %>% filter(cohort < 2020 | is.na(cohort) == T) %>% group_by(cohort, inst) %>% summarise(n = n_distinct(ID)) %>% spread(inst, n) # 2103 all students records
data %>% filter(cohort < 2020 | is.na(cohort) == T) %>% filter(Graduated. == "Y") %>% group_by(cohort, inst) %>% summarise(n = n_distinct(ID)) %>% spread(inst, n) # 241 (11.4%)
two_year_cohort <- data %>% filter(Graduated. == "Y") %>% filter(Graduation.Date < twoyear) %>% group_by(cohort, inst) %>% summarise(n = n_distinct(ID)) %>% spread(inst, n) #138 (6.5%)
four_year_cohort <- data %>% filter(Graduated. == "Y") %>% filter(Graduation.Date < fouryear) %>% group_by(cohort, inst) %>% summarise(n = n_distinct(ID)) %>% spread(inst, n) #180 (8.5%)
six_year_cohort <- data %>% filter(Graduated. == "Y") %>% filter(Graduation.Date < sixyear) %>% group_by(cohort, inst) %>% summarise(n = n_distinct(ID)) %>% spread(inst, n) # 213 (10.1)
```

```{r}
knitr::kable(all_student_cohort, caption = "All student by cohort by institution")
```


```{r}
knitr::kable(two_year_cohort, caption = "Cohort graduation at two years", col.names = c("Cohort", "Davis Tech", "Mtn. Land"))
```

```{r}
knitr::kable(four_year_cohort, caption = "Cohort graduation at four years", col.names = c("Cohort", "Davis Tech", "Mtn. Land"))
```

```{r}
knitr::kable(six_year_cohort, caption = "Cohort graduation at six years", col.names = c("Cohort", "Davis Tech", "Mtn. Land"))
```

\newpage

# SLCC transfers and graduations

```{r}
two_year_trans <- data %>%
  filter(College.Name == "SALT LAKE COMMUNITY COLLEGE") %>%
  filter(StartDate <= twoyear) %>% group_by(cohort, inst) %>%
  summarise(n = n_distinct(ID)) %>%
  spread(inst, n) #138
two_year_trans[is.na(two_year_trans) == T] <-0

knitr::kable(two_year_trans[1:9,], 
             caption = "Two year transfer to SLCC by cohort year.",
             col.names = c("Cohort", "DTC", "MTL"))
```

```{r}
six_year_cohort_slcc <- data %>% 
  filter(Graduated. == "Y") %>%
  filter(College.Name == "SALT LAKE COMMUNITY COLLEGE") %>% 
  filter(Graduation.Date < sixyear) %>% 
  group_by(cohort, inst) %>%
  summarise(n = n_distinct(ID)) %>%
  spread(inst, n) #138 (6.5%)

knitr::kable(six_year_cohort_slcc[1:6,],
             caption = "Six year graduations at SLCC by cohort year.",
             col.names = c("Cohort", "DTC"))
```


